Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching

### 2025-09-26 review_responses.tex

\usepackage{graphicx}   % for \includegraphics
\usepackage{caption}    % better caption control inside minipage
\usepackage{subcaption} % optional, but useful if you want subfigures later

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  urlcolor=blue!60!black,
  citecolor=blue!60!black
}
\setcounter{secnumdepth}{1}
% Compact lists
\setlist[itemize]{leftmargin=1.2em,itemsep=0.3em,topsep=0.3em}
\setlist[enumerate]{leftmargin=1.4em,itemsep=0.3em,topsep=0.3em}

% Handy macros used in your text
\newcommand{\BPNoise}{\textsc{BP-Noise}}
\newcommand{\AmpMod}{\textsc{AmpMod}}
\newcommand{\LossSelect}{\textbf{Loss Selection}}
\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
\newcommand{\PeriodicLoss}{\textbf{Loss Landscape Navigation}}
\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
\newcommand{\DTWEnv}{\textsc{DTW\_Envelope}}
\newcommand{\SIMSESpec}{\textsc{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\textsc{L1\_Spec}}
\newcommand{\JTFS}{\textsc{JTFS}}

\title{\vspace{-0.5em}Author Response for T-ASL-11868-2025\\
\large Manuscript: \emph{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}}
\author{}
\date{}

@@

\begin{document}
\maketitle
\vspace{-1.25em}

\noindent Dear Editor Alberto Bernardini, \\
\noindent
We'd like to thank you and the reviewers for the thoughtful and constructive feedback on our manuscript. We have carefully considered all comments and revised the manuscript accordingly. Here you will find our detailed point-by-point response to the feedback.

For each reviewer, we first provide the review in full, with the key concerns numbered and in bold. Below the full review, we address the concerns in dedicated sections. Each section contains the reviewer's key requests or issues, followed by our response. Some revisions address more general concerns expressed in the body of the reviews (\nameref{R1.1} and \nameref{R1.2} for example). As such, for the sake of brevity we took the liberty to paraphrase the reviewer's requests and concerns as accurately as possible when necessary.

@@

\medskip
\noindent
We hope the revisions address all concerns and improve the clarity and quality of the manuscript.

\noindent Sincerely, Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane

\section{Reviewer 1}
\subsection{Original Review}

@@

\noindent To further clarify the methodology and its importance, we have added a new subsection, \textbf{II.H What Approach Should We Take?} placed immediately before the Methodology section. This subsection clarifies our experimental design and explains why our contribution is unique. Specifically, we use:  
\begin{enumerate}
  \item Simple differentiable DSP synthesizers to isolate synthesis--loss interactions,  
  \item Multiple losses tested side by side within the same framework, and  
  \item Direct gradient descent optimization without neural proxies.  
\end{enumerate}

This quote from the aforementioned section in the paper makes clear that our work is not simply a report of evaluations, but a controlled benchmarking study designed to reveal how losses and synthesis methods interact:

\begin{displayquote}
    We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems. To address the problems of \LossSelect{} and \SynthSelect{}, we adopt a methodology with three key characteristics that have not previously appeared together.

\begin{enumerate}
    \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem. Restricting to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions. 
    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem. Prior work has typically tested losses in isolation, making direct comparisons rare\cite{vahidi2023mesostructures}.  
    \item Optimize synthesizer parameters directly with gradient descent, without neural approximations. This complements prior proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures. 
\end{enumerate}

@@

\subsection*{R2.3 Clear takeaways}
\noindent\textbf{Review:}
\textbf{This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.\\}

\noindent\textbf{Response:} \\
In the revised paper, we've added two sections which address the valid issues raised here and by Reviewer 1.

\begin{itemize}
  \item We added a new subsection, \emph{Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
  \item We introduced a new subsection, \emph{Practical Recommendations}, which distills our findings into actionable heuristics for practitioners. DTWEnv is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised). We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.

\end{itemize}

For a more detailed response, see our response to the Reviewer.1 in Sections~\nameref{R1.6} and \nameref{R1.7}.

\bibliographystyle{IEEEtran}
\bibliography{references}

### 2025-09-26 review_responses.tex

\title{\vspace{-0.5em}Author Response for T-ASL-11868-2025\\
\large Manuscript: \emph{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}}
\author{}
\date{}

\begin{document}
\maketitle
\vspace{-1.25em}

\noindent Dear Editor Alberto Bernardini, \\
\noindent
We'd like to thank you and the reviewers for the thoughtful and constructive feedback on our manuscript. We have carefully considered all comments and revised the manuscript accordingly. Here you will find our detailed point-by-point response to the feedback.

For each reviewer, we first provide the review in full, with the key concerns numbered and in bold. Below the full review, we address the concerns in dedicated sections. Each section contains the reviewer's key requests or issues, followed by our response. Some revisions address more general concerns expressed in the body of the reviews (\nameref{R1.1} and \nameref{R1.2} for example). As such, for the sake of brevity we took the liberty to paraphrase the reviewer's requests and concerns as accurately as possible when necessary. We hope these revisions have addressed your concerns and improved the clarity and quality of the manuscript.

\medskip
\noindent Sincerely, Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane

@@

  \item \textbf{(R1.3) Restructure the manuscript to clarify the boundary between methodology, results, and discussion.}
\end{enumerate}

\subsection*{R1.1: Introduction Clarity}
\label{R1.1}
\noindent\textbf{Reviewer:}\\
\textbf{We recommend tightening the Introduction to focus more directly on problem motivation and contribution, and removing or completing the subsection structure.
\\}

\noindent\textbf{Response:} \\
We agree with the reviewer’s comment regarding lack of clarity and focus in the Introduction. The Introduction has been revised accordingly: 
\begin{itemize}
  \item The Introduction is now a single continuous section, without subsection headings.  
  \item We have removed unnecessary historical context and repeated definitions, reducing the length by roughly 25\%. We moved some of this material to the new \textbf{II. Background and Related Work} section.  
  \item The central hypothesis now appears at the beginning of the third paragraph (previously it appeared mid-way through paragraph 10).  
  \item In addition to the main hypothesis, we have added three explicit research questions  (answered in \textbf{IV.B Practical Recommendations}), covering metric agreement (Q1), the utility of DTW and SIMSE in implementing novel losses (Q2), and the viability of iterative optimization (Q3). These are stated directly after the central hypothesis:

@@

  \item The Introduction concludes with a concise statement of our contributions:

  \begin{displayquote}
      The contributions of this paper include (1) an evaluation of multiple differentiable losses across multiple synthesis methods, (2) the introduction and justification of loss functions utilizing DTW and SIMSE (3) a discussion on the utility and agreement between manual and automatic evaluation metrics (4) a nomenclature of different sound-matching approaches and unsolved issues in the field. 
  \end{displayquote}
\end{itemize}

\subsection{R1.2 Improve Background and Related Work}
\label{R1.2}
\noindent\textbf{Reviewer:}\\
\noindent\textbf{The current division of the manuscript into Introduction, Background, and Past Works results in significant conceptual overlap and repetition. Historical framing, loss function taxonomy, and method classification are spread across all three sections. A clearer organizational approach would be to merge Background, and Past Works into a unified Background and Related Work section, where formal definitions, theoretical framing, and literature review are presented coherently. This would allow the Introduction to focus solely on the motivation, problem framing, and contribution.}\\

\noindent\textbf{Response:} \\
We agree that the organization and writing in these sections could be improved. We merged \emph{Background} and \emph{Past Works} into a unified section, \textbf{II. Background and Related Work}. We applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions in compromising the flow.
\textbf{II. Background and Related Work} contains the following topics in order:  
\begin{itemize}
  \item A formal definition of sound-matching and its components.
  \item Definitions of synthesizers, losses, and optimization heuristics. 
  \item A chronological review of prior sound-matching approaches.
  \item A discussion of gaps in the field and four critical issues.
  \item A justification for our methodology, and how our experiments address the two critical issues of \emph{Loss and Synthesizer Selection}.
\end{itemize}

\subsection*{R1.3 Organize Methodology, Results, and Discussion.}
\label{R1.3}
\noindent\textbf{Reviewer:}\\
\textbf{Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).}
\\

@@

\noindent\textbf{Response:} \\
We agree that the earlier draft blurred the boundaries between setup, results, and discussion. In the revised manuscript we implemented the reviewer’s suggestion as follows:

\begin{itemize}
  \item \textbf{III. Experimental Setup and Results A-E}: Sections III.A–E now exclusively describe the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests. Each subsection is focused on describing how the experiments are designed, without interpretation of outcomes. For instance, the synthesizer section defines parameter ranges, while the evaluation section specifies how statistical significance is tested. The only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program}, which serves as a transition between results and discussion.

@@

  \item Experimental findings are presented in \textbf{III-F Best Performers for Each Program} and \textbf{III-G Consistency in Rankings}, which together form the \emph{Results} component. Section III-F reports which losses performed best for each synthesizer, while Section III-G quantifies the consistency of rankings across bootstrap resamples. Keeping all numerical findings here avoids overlap with later interpretation.

  \item \textbf{IV. Discussion} then provides the qualitative interpretation, structured into four dedicated subsections:  
    \begin{enumerate}
      \item \textbf{IV-A. Key Findings}: summarizes the central empirical trends, such as the dependency of loss performance on synthesis method.  
      \item \textbf{IV-B. Practical Recommendations}: translates results into concrete recommendations for practitioners and researchers.  
      \item \textbf{IV-C. Illustrating Loss Landscapes}: adds visualizations that explain why certain losses succeed or fail in specific contexts.  
      \item \textbf{IV-D. Applicability to More Complex Synthesizers}: discussion on how observed patterns extend beyond the simplified testbed to richer, real-world synthesizers.  
    \end{enumerate}

@@

  \item To improve clarity, we also relocated explanatory material about experiment design (e.g., parameter settings, optimizer configuration) previously scattered across the results section into Sections III-A–E, so that the results sections are strictly empirical while the discussion is interpretive.  
\end{itemize}

Overall, this restructuring produces a clearer separation of topics: methodology is isolated in Section III-A–E, empirical findings are concentrated in Section III-F–G, and broader interpretation is reserved for Section IV. We believe this major restructuring and rewrites of appropriate sections has improved flow of the manuscript and made the boundary between methodology, results, and discussion much clearer.

\subsection*{R1.4 Justification for \DTWEnv{} and \SIMSESpec}
\label{R1.4}
\noindent\textbf{Reviewer:} \\
\noindent While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification.

\noindent\textbf{Response:} \\
We have expanded the theoretical justifications for DTW\_Envelope and \SIMSESpec{} in subsection \textbf{II.C Sound Representation and Loss Functions}. In \textbf{II.C.2 Fourier Spectrograms}, we discuss the possible application of SIMSE as an alternative to L1:
\begin{displayquote}
    A common limitation of spectrogram-based losses is their sensitivity to global gain: two signals with identical spectral shape but different overall amplitude can yield large errors under L1/L2 norms. 
    Here we propose and test Scale-Invariant Mean Squared Error (SIMSE) as an alternative to L1/L2 norms in spectrogram comparisons. SIMSE normalizes the amplitudes before comparison and emphasizes proportional differences in spectral envelopes~\cite{barron2014shapessimse}. 
    Perceptually, listeners are often tolerant of loudness changes while remaining sensitive to timbral shape, therefore this property could be advantageous for subtractive synthesis, where filter cutoffs reshape spectral balance without predictable changes in total energy. Although SIMSE has been applied in other contexts such as audio reconstruction, to our knowledge its use as a differentiable spectrogram loss in iterative sound-matching is novel.
\end{displayquote}

 In \textbf{II.C.4 Dynamic Time Warping} For DTW\_Envelope, we clarify its perceptual motivation: aligning amplitude envelopes is crucial for AM synthesis, where traditional spectrogram distances fail to capture similarity: 
 \begin{displayquote}
     DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment. 
    For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
    A possible use case to consider is the application of DTW to amplitude envelopes, which may be able to match two sounds with similar rates of loudness modulation regardless of alignments. 
    To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).
 \end{displayquote}

 We explicitly frame these losses not as wholly new techniques, but as carefully adapted methods to be tested in an iterative differentiable setting, a contribution that has not previously been made.

@@

We did not modify the definitions in \textbf{III.A Loss Function Implementation Details}, dedicating this section to only technical explanations.

\subsection*{R1.5 Analysis of Failure or Success.}
\label{R1.5}
\noindent\textbf{Reviewer:} \\
\textbf{Add qualitative or spectral analyses of where specific losses succeed or fail}\\

\noindent\textbf{Response:} \\
This is an interesting suggestion. Our original experiments were designed to emphasize quantitative, statistically robust measurements of loss function performance, which is why we initially avoided focusing on individual examples of sounds, spectrograms, or loss landscapes (in part to reduce the risk of cherry-picking). We do concede that illustrative examples can provide helpful intuition for readers and motivate future research, however, we believe that an analysis of loss landscapes (rather than spectrograms or qualitative tests) is a more useful approach to an explanation for \textit{where} and \textit{why} losses might succeed or fail.

To this end, we have added a new Section \textbf{IV.C llustrating Loss Landscapes} where qualitative visualizations of loss landscapes  are added. These examples highlight why specific losses may succeed or fail and reinforce the broader importance of loss landscapes in assessing the appropriateness of similarity measures~\cite{vahidi2023mesostructures,turian2020sorry}, supporting our argument that effective navigation of loss landscapes represents an important gap in current knowledge and a possible direction for future work:

\begin{displayquote}
We show the landscapes of the loss functions with simplified versions of \BPNoise{} (HP-Noise) and \AmpMod{} (S-Noise-AM), each reduced to a single parameter for one-dimensional visualization. HP-Noise uses only a high-pass filter with cutoff between 100–20,000 Hz, while S-Noise-AM varies only the modulation rate between 0.1–20 Hz. Loss values are normalized to 0–1 to allow comparison across functions.

    Figure~\ref{fig:loss_landscape_noisebp} [\textbf{Figure 4 in paper}] shows the HP-Noise results. \LoneSpec{}, \SIMSESpec{}, and \JTFS{} exhibit clear minima at the correct parameter (red dashed line), whereas \DTWEnv{} remains relatively flat, making gradient-based optimization more difficult.

    Figure~\ref{fig:loss_landscape_ampmod} [\textbf{Figure 5 in paper}]  shows the S-Noise-AM results. Here, \DTWEnv{} produces the smoothest and most informative landscape, while spectrogram losses are minimized near the target but lack consistent correlation with parameter distance. \JTFS{} remains largely flat. This aligns with \DTWEnv's superior performance in amplitude-modulated synthesis observed in our experiments.
\end{displayquote}

@@

\begin{figure*}[ht]
    \centering
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_loss_landscapes_normalized.png}
        \caption{Loss landscapes for \BPNoise{} with only a high-pass filter parameter. 
        \LoneSpec{}, \SIMSESpec{}, and \JTFS{} show clear global minima near the correct parameter, while \DTWEnv{} remains flat around the target.}
        \label{fig:loss_landscape_noisebp}
    \end{minipage}%
    \hfill
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_lanscapes_1_1d_normalized.png}
        \caption{Loss landscapes for a simplified \AmpMod{} synthesizer with only amplitude modulation parameter. 
        \DTWEnv{} exhibits the smoothest and most informative landscape, explaining its superior performance in amplitude-modulated synthesis.}
        \label{fig:loss_landscape_ampmod}
    \end{minipage}
\end{figure*}

While this may not be the "qualitative" or "spectral" analysis, we believe it is instructive on the mechanisms underlying our quantitative findings and highlight the importance of loss landscape navigation as an open area for future research.

\subsection*{R1.6 Applicability to More Complex or Real-world Synthesizers.}
\label{R1.6}
\noindent\textbf{Review:} \\
\noindent \textbf{Expand the discussion to address applicability to more complex or real-world synthesizers}
\\

@@

\begin{displayquote}
    The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principle with 2 parameters each. While this level of simplicity facilitates controlled comparisons, it does not fully capture the richness of real-world synthesizers, which can have hundreds of parameters with various routes. As a result, the extent of the generalizability of our findings does require further research.

    Nonetheless, we believe that the insights here can extend naturally to more complex domains. Depending on the feature of interest, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.

    \DTWEnv{} measures very specific features of audio, and its use in the recommended settings---where low frequency amplitude modulation matching is required---would likely yield the desired results regardless of synthesizer complexity. Likewise, the use of \SIMSESpec{} or \LoneSpec{} for matching filter cut-offs is likely to be successful regardless of the underlying sound. 
    \end{displayquote}

\subsection{R1.7 Provide Practical Takeaways}
\label{R1.7}
\noindent\textbf{Review:} \\
\noindent \textbf{Provide practical takeaways or usage guidelines for different loss-synth combinations. } 
\\

\noindent\textbf{Response} \\

In \textbf{IV.A Key Findings}, we discuss important findings and provide answers to our research questions. The following section, \textbf{IV.B Practical Recommendations}, distills our findings into actionable heuristics for practitioners.

\begin{displayquote}
    Our findings not only provide answers to the research questions proposed in the introduction, but also serve as guidelines for practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
\begin{itemize}
    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking. However, we encourage confirmation of conclusions with manual listening, particularly if specificity is important. 
    \item For amplitude-modulated synthesis, DTW-based losses such as \DTWEnv{} are recommended over spectrogram differences and JTFS.
    \item  For subtractive/noise-filtered synthesis, spectrogram based losses are most effective. Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
    \item Iterative differentiable optimization via the Faust-to-JAX pipeline is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
    \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}. Thus, our results call into question its effectiveness as a general mesostructure measure, and highlight a need for further research into its utility.
\end{itemize}

@@

\subsection{R2.1 Clarity of Proposed Method}
\label{R2.1}
\noindent\textbf{Review:}
\noindent \textbf{It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations. } 
\\

\noindent\textbf{Response:} \\
We have taken several measures to emphasize the importance and reasoning behind our methodology design and ensure that the paper is not read as a set of evaluation results.

\medskip
\noindent In \textbf{II.B Digital Signal Processing and Synthesis}, we state the importance of isolated sound-matching experiments for analysis of loss-synthesizer interactions:
\begin{displayquote}

@@

\noindent To further clarify the methodology and its importance, we have added a new subsection, \textbf{II.H What Approach Should We Take?} placed immediately before the Methodology section. This subsection clarifies our experimental design and explains why our contribution is unique. Specifically, we use:  
\begin{enumerate}
  \item Simple differentiable DSP synthesizers to isolate synthesis--loss interactions,  
  \item Multiple losses tested side by side within the same framework, and  
  \item Direct gradient descent optimization without neural proxies.  
\end{enumerate}

This quote from the aforementioned section in the paper makes clear that our work is not simply a report of evaluations, but a controlled benchmarking study designed to reveal how losses and synthesis methods interact:

\begin{displayquote}
    We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems. To address the problems of \LossSelect{} and \SynthSelect{}, we adopt a methodology with three key characteristics that have not previously appeared together.

\begin{enumerate}
    \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem. Restriction to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions. 
    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem. Prior work has typically tested losses in isolation, making direct comparisons of loss functions rare~\cite{vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}, particularly if comparisons are done in the context of different methods of synthesis. 
    \item Optimize synthesizer parameters directly with gradient descent, without neural approximations. This complements prior proxy-based methods by providing the missing “low-level experiments” that aim to analyze fundamental interactions between DSP functions and similarity measures. 
\end{enumerate}

@@

\end{displayquote}

\medskip
\noindent We restate our methodology in the opening paragraph of the \textbf{III. Experimental Setup} section:

### 2025-09-26 sections/background_related.tex
Updates from Overleaf

\subsection{Formalization of Sound-Matching}
\label{sec:sound_matching_definition}
Following prior works~\cite{vahidi2023mesostructures,han2023perceptual}, we define sound-matching in terms of a parametric synthesizer, a target sound, a representation function, and a similarity measure. 
The key components are:
\begin{itemize}
    \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
    \item $x$: The synthesizer output, $x = g(\theta)$.
    \item $t$: The target sound to be replicated or imitated. 
    \item $\phi(\cdot)$: A representation (feature extraction) function mapping signals into a comparison space.
    \item $L$: A loss function that measures distance between $x$ and $t$, typically via $L(\theta,t) = d(\phi(x), \phi(t))$ for some metric $d$.
\end{itemize}

@@

    % Nodes
    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};

@@

\[
L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
\]

@@

A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
The proportionality or distance metric $d$ has typically been the L1 or L2 distance~\cite{turian2020sorry,richard2025model}, with L1 being calculated as the mean of the absolute difference between every point in the proxy representation:
\[
L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
\]

@@

DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment. 
For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
By applying DTW specifically to amplitude envelopes, we can match two sounds with identical rates ofloudness modulation regardless of alignments. 
To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).

\subsection{In-Domain versus Out-of-Domain}
\label{sec:in-domain}
The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, the scenario that is called \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}. To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}. If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}. In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound. If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.

@@

\subsection{Supvervised vs Direct Optimization}
\label{sec:optimization}

The choice of \textit{heuristics} is yet another important attribute sound-matching. The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
\[
\theta^* = \arg\min_{\theta} L(\theta,t)
\]

@@

Differentiable loss functions that use spectrogram differences can be computationally expensive. To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP). PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function. PNP loss functions are fast and differentiable, but require training and parameter estimation. A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}. 
\[
\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
\]

@@

A neural approximation for another part of the supervised sound-matching chain---this time the synthesizer---was proposed by Barkan \textit{et al.}~\cite{barkan2023inversynthII}. As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $x_o$ to $x_t$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}. This approach used a model mapping sounds to parameters, another approximating the synthesizer, and a loss function which combines P-Loss with STFT differences. This ``Inversynth II'' (IS2) approach yielded significant improvements to previous works which did not use the STFT approximations for loss~\cite{esling2019flow,barkan2019inversynth}. Barkan \textit{et al.} then attempted to improve the IS2 model with Inference-Time
Fine-tuning (ITF). For ITF, the synthesizer approximation is frozen and the encoder is iteratively fine-tuned for a particular sample. However, ITF often degraded performance, possibly due to proxy–synthesizer mismatch or overfitting.

@@

\begin{enumerate}
    \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem. Restricting to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions. 
    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem. Prior work has typically tested losses in isolation, making direct comparisons rare~\todo{cite}.  
    \item Optimize synthesizer parameters directly with gradient descent, without neural approximations. This complements prior proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures. 
\end{enumerate}

>>>>

\subsection{Formalization of Sound-Matching}
\label{sec:sound_matching_definition}
Following prior works~\cite{vahidi2023mesostructures,han2023perceptual}, we define sound-matching in terms of a parametric synthesizer, a target sound, a representation function, and a similarity measure. 
The key components are:
\begin{itemize}
    \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$.”.
    \item $\xzero$: The target sound to be replicated or imitated. 
    \item $\phi(\cdot)$: A representation (feature extraction) function mapping signals into a comparison space.
    \item $L$: A loss function that measures distance between $\xtheta$ and $\xzero$, typically via $L(\theta,\xzero) = d(\phi(\xtheta), \phi(\xzero))$ for some metric $d$.
\end{itemize}

@@

% Nodes
\node (start) [text centered] {\( g(\hat{\theta}) = x_{\theta} \)};
\node (L) [above of=start, right of=start, text centered] 
    {\( L(\hat{\theta}, x_{0}) \propto \phi(x_{\theta}) - \phi(x_{0}) \)};
\node (optimize) [below of=L, right of=L, text centered] {Optimize $\hat{\theta}$};
\node (new_theta) [below of=optimize, right of=start, text centered] {New \( \hat{\theta} \)};

@@

\[
L(\theta, \xzero) = d(\phi(\xtheta), \phi(\xzero))
\]

@@

A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds  $\xzero$ and $\xtheta$  to their respective representations. 
The proportionality or distance metric $d$ has typically been the L1 or L2 distance~\cite{turian2020sorry,richard2025model}, with L1 being calculated as the mean of the absolute difference between every point in the proxy representation:
\[
L(\theta, \xzero) = \left\| \phi(\xzero) - \phi(\xtheta) \right\|_1
\]

@@

DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment. 
For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
A possible use case to consider is the application of DTW to amplitude envelopes, which may be able to match two sounds with similar rates of loudness modulation regardless of alignments. 
To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).

\subsection{In-Domain versus Out-of-Domain}
\label{sec:in-domain}
The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, the scenario that is called \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}. To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $\xzero$, or put differently, if $\xzero$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}. If $\xzero$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}. In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound. If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.

@@

\subsection{Supvervised versus Direct Optimization}
\label{sec:optimization}

The choice of \textit{heuristics} is yet another important attribute sound-matching. The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
\[
\theta^* = \arg\min_{\theta} L(\theta,\xzero)
\]

@@

Differentiable loss functions that use spectrogram differences can be computationally expensive. To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP). PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function. PNP loss functions are fast and differentiable, but require training and parameter estimation. A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.

\[
\| \varphi(\xzero) - \varphi(\xtheta) \|_2^2
= \langle \tilde{\theta} - \theta \,|\, M(\theta) \,|\, \tilde{\theta} - \theta \rangle
+ O(\|\tilde{\theta} - \theta\|_3^2).
\]

@@

A neural approximation for another part of the supervised sound-matching chain---this time the synthesizer---was proposed by Barkan \textit{et al.}~\cite{barkan2023inversynthII}. As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $\xtheta$ to $\xzero$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}. This approach used a model mapping sounds to parameters, another approximating the synthesizer, and a loss function which combines P-Loss with STFT differences. This ``Inversynth II'' (IS2) approach yielded significant improvements to previous works which did not use the STFT approximations for loss~\cite{esling2019flow,barkan2019inversynth}. Barkan \textit{et al.} then attempted to improve the IS2 model with Inference-Time
Fine-tuning (ITF). For ITF, the synthesizer approximation is frozen and the encoder is iteratively fine-tuned for a particular sample. However, ITF often degraded performance, possibly due to proxy–synthesizer mismatch or overfitting.

@@

\begin{enumerate}
    \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem. Restriction to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions. 
    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem. Prior work has typically tested losses in isolation, making direct comparisons of loss functions difficult~\cite{vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}. Particularly rare are comparisons which are done in the context of different methods of synthesis.  
    \item Optimize synthesizer parameters directly with gradient descent, without neural approximations. This complements prior proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures. 
\end{enumerate}
